---
title: "SqTSq (process)"
author: "George Ionitsa"
---

```{r}
library(lpSolve)
```

This code helps to make all possible 'squares' in a matrix 5x5. 
VM will be the list of all matrices. In every new iteration we add an additional matrix in the list.

I do not like how I create zero M matrix every time. I believe it can be done more beautiful.

Overall, this is a code with 3 whiles, where there the first one is a loop for different sizes of squares and the next two are simple loops to fill ones in the right order. 

```{r}
VM <- list()
count <- 1
M <- matrix(0, nrow = 5 , ncol = 5) #creating 5x5 zero matrix
i = 1
j = 1
size = 0 
while (size <= 3){
  while (i <= 5 - size){
    while (j <= 5 - size){
      M <- matrix(0, nrow = 5 , ncol = 5) #reset this matrix every time
      M[i:(i+size),j:(j+size)] <- 1
      print(M)
      j = j+1
      VM[[count]] <- M
      count = count + 1
    }
    j = 1
    i = i+1
  }
j = 1
i = 1  
size = size + 1
}
```

Next step is to represent matrices as strings. Firstly, I started to write it purely mathematical. However, I realised that this is R and probably there is a special function for this and voila: ```as.vector()```.


```{r}
HM <- c() #HM stands for Huge Matrix since it will be huge. 
count <- 1
while (count <= 54){
  HM <- rbind(HM,as.vector(VM[[count]])) #simply use rbind to connect arrays 
  count = count + 1 
} 
HM <- t(HM) #transpose to bring it to the right dimensions
print(HM)
```

Now we can apply lp to solve this problem.

```{r}
f.obj <- rep(1,54)
f.con <- HM
f.dir <- rep("=", 25)
f.rhs <- rep(1, 25)
lp("min", f.obj, f.con, f.dir, f.rhs)$solution
```


This is great that our code gave us integer positive solutions. 
However, I think it is neccessary to add additional constraints in terms of the form: xi <= 1 for each i. 

I assume it can be done like this:

```{r}
f.obj <- rep(1,54)
f.con <- rbind(HM,diag(54))
f.dir <- c(rep("=", 25),rep("<=", 54))
f.rhs <- c(rep(1, 25), rep(1,54))
x <- lp("min", f.obj, f.con, f.dir, f.rhs)$solution
```

I hope it does actually work like this. 
Now I will try to beautifully draw it. 
I have no clue how to do it more efficient without having 4 consecutive whiles:
(I know how. It is one while function with an idea of a sum of consecutive squares.)

```{r}
RM <- matrix(0, nrow = 5, ncol = 5) #RM stands for Result Matrix since it will be a result. 
count <- 1
while (count <= 25){
  if (x[count] == 1)
    RM <- RM + VM[[count]]  
    count = count + 1 
} 

while (count <= 41){
  if (x[count] == 1)
    RM <- RM + 2*VM[[count]]  
    count = count + 1 
} 


while (count <= 50){
  if (x[count] == 1)
    RM <- RM + 3*VM[[count]]  
    count = count + 1 
} 


while (count <= 54){
  if (x[count] == 1)
    RM <- RM + 4*VM[[count]]  
    count = count + 1 
} 
print(RM)
```

Now we can distinguish 1x1, 2x2, 3x3 and 4x4 squares. For example, if there is 2 in a matrix - it is a part of a 2x2 square.

Since my code with 3 whiles is pretty universal, it is very simple to upgrade it to the n-case.  
All we need to do is to change ```size``` and ```M <- matrix(0, nrow = n , ncol = n)```
I hope it will work.

Comment: I think I found better way to reset matrix. Instead of creating a zero matrix again, I put zeros back in given cells.

```{r}
n = 7
VM <- list()
count <- 1
M <- matrix(0, nrow = n , ncol = n) #creating 5x5 zero matrix
i = 1
j = 1
size = 0 
while (size <= n-2){
  while (i <= n - size){
    while (j <= n - size){
      M <- matrix(0, nrow = n , ncol = n) 
      M[i:(i+size),j:(j+size)] <- 1
      j = j+1
      VM[[count]] <- M
      count = count + 1
    }
    j = 1
    i = i+1
  }
j = 1
i = 1  
size = size + 1
}
```

I am surprised, but it worked. However, my idea with having ```M[i:(i+size),j:(j+size)] <- 1``` instead of ```M <- matrix(0, nrow = n , ncol = n) ``` failed. 
I probably now why. Because, I put this ```M[i:(i+size),j:(j+size)] <- 0``` after I change value of j. Now it is fixed. 
Second try:

```{r}
n = 7
VM <- list()
count <- 1
M <- matrix(0, nrow = n , ncol = n) #creating 7x7 zero matrix
i = 1
j = 1
size = 0 
while (size <= n-2){
  while (i <= n - size){
    while (j <= n - size){
      M[i:(i+size),j:(j+size)] <- 1
      VM[[count]] <- M
      M[i:(i+size),j:(j+size)] <- 0
      count = count + 1
      j = j+1
    }
    j = 1
    i = i+1
  }
j = 1
i = 1  
size = size + 1
}
```

Ok.

```{r}

check = n
amofel = 0
while (check > 1){
  amofel = amofel + check*check
  check = check-1
} 

##It is to understand how many possible squares we will have.  

```

```{r}
HM <- c() #HM stands for Huge Matrix since it will be huge. 
count <- 1
while (count <= amofel){
  HM <- rbind(HM,as.vector(VM[[count]])) #simply use rbind to connect arrays 
  count = count + 1 
} 
HM <- t(HM) #transpose to bring it to the right dimensions

```

```{r}
f.obj <- rep(1, amofel)
f.con <- HM
f.dir <- rep("=", n*n)
f.rhs <- rep(1, n*n)
solution <- lp("min", f.obj, f.con, f.dir, f.rhs, all.int=TRUE)$solution
print(solution)

```

Many apologises for a lack of comments in a last part. However, it is done for n-case.

```{r}
i = 1
count = 0
while (i <= length(solution)){
  if (solution[i] == 1){
    print(VM[[i]])
    count = count + 1
  }
  i = i+1
}

print(count)
```